\Wxxsir methods to estimate $\rr$ in the SIR model

### 2018-07-03 paper/Version13/_region_.tex
finally on git

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
\usepackage{systeme}
%% tables
\usepackage{booktabs}
\usepackage{lscape}
\usepackage[table,xcdraw]{xcolor}
\usepackage{float}

@@

\newcommand{\xxsir}{\ensuremath{9}} % number of SIR methods
\newcommand{\wxxsir}{nine }
\newcommand{\Wxxsir}{Nine }

@@

@@

In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.  This method assumes that we have some prior knowledge about $\gamma$, the transmission rate.  Thus we use as our estimate,
\begin{align*}
\hat{\rr} &= \textnormal{IPR}(t) \cdot \frac{1}{\gamma}
\end{align*}

@@

\subsection{Exponential Growth}\label{sec:expgrowth}
%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.

@@

We compare the performance of the SIR models on simulated data from both the SIR as well as other models. In all of these simulations, we first generate data from the models under known conditions, then we add on errors with known variances to each time point,
\begin{align}\label{eq:sim-models}
  X_{model}(t) &= f(t) + \epsilon_{X,t} \\
  Y_{model}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
  Z_{model}(t) &= N - X(t) - Y(t)\nonumber 
\end{align}
We adjust the recovered compartment $Z$ so the total population is constant. 
We simulate with two different types of error distributions, Gaussian and autoregressive,

@@

\noindent \textbf{Gaussian-PAVA (Norm-M) ($X_{GP}$, $Y_{GP}$, $Z_{GP}$)}
\begin{align*}
 X_{GP}(t) &= \textnormal{PAVA}(X_G(t)) \\
  Y_{GP}(t) &= N - X_{GP}(t) - Z_{GP}(t) \\
  Z_{GP}(t) &= \textnormal{PAVA}(Z_G(t)) \\
\end{align*}
\textbf{Autoregressive-PAVA (AR-M) ($X_{ARP}$, $Y_{ARP}$, $Z_{ARP}$)}
\begin{align*}
  X_{ARP}(t) &= \textnormal{PAVA}(X_{AR}(t)) \\
  Y_{ARP}(t) &= N - X_{ARP}(t) - Z_{ARP}(t) \\
  Z_{ARP}(t) &= \textnormal{PAVA}(Z_{AR}(t)) 
\end{align*}
This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M) for a given set of parameters $(\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N)$.

@@

We first simulate 5 ``baseline'' data sets from the SIR model under the following conditions:

@@

	Total population $N$: 100000.

@@

We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison. The different data sets generate are shown in the following table:

\begin{table}[h]
	\centering
	\resizebox{\textwidth}{!}{
		\begin{tabular}{@{}lll@{}}
			\toprule
                  \textbf{Parameter Values ($\beta, \gamma$)}        & \textbf{Starting Compartment Sizes ($X_0, Y_0$)}            & \textbf{Variance of Added Errors ($\sigma_X, \sigma_Y$ )}       \\ \midrule
                  (0.06, 0.001)            & (99000, 1000)                           & (50, 2)                     \\
                  (0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
                  (0.06,.03) & (99950, 50) & (100, 5)\\
                  (0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
                  (0.06, 0.24) & (99999, 1) & (10000, 500)                                      \\ \bottomrule
		\end{tabular}
	}
	\caption{Table of data sets.}
	\label{tab:datasets}
\end{table}

@@

\section{Results}\label{sec:results}
There are numerous aspects to consider when using a method to estimate $\rr$.  Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available (time $T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, and  the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X, \sigma_Y$).  We examine each of these aspects individually.

@@

\begin{table}[H]
	\includegraphics[scale=0.5]{BaseBase.jpeg}

	\caption{\label{tab:}Baseline Data, $\rr$ Estimates and Standard Errors}
	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		SIRLS & 1.9999711 & 0.0055923\\
		\hline
		SIRREPAR & 1.9997340 & 0.0049939\\
		\hline
		SIRLMA & 1.6863640 & 0.1886132\\
		\hline
		SIRLMAT & 1.4063227 & 0.6309170\\
		\hline
		SIRIPR & 4.4203632 & 12.3593415\\
		\hline
		SIRSIPR & 1.8870824 & $<$ 1e-07 \\
		\hline
		LOGLINEAR & 1.9998072 & 0.0002198\\
		\hline
		SIRMC & 0.9485555 &  $<$ 1e-07 \\
		\hline
		SEQBAYES & 1.6495023 & 0.0672249\\
		\hline
	\end{tabular}
\end{table}

@@

\includegraphics[scale=0.5]{AR.jpeg}
\begin{table}[H]

@@

\textbf{Need to figure out incorporating different $/rr$ scales in to graph. Facets won't work.}

@@

\includegraphics[scale=0.5]{time.jpeg}

@@

\includegraphics[scale=0.5]{start.jpeg}

@@

\includegraphics[scale=0.5]{var.jpeg}

@@

  \begin{center}
\includegraphics[scale=0.5]{other.jpeg}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLS Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline1 & 1.999919 & 0.0058322 & 2.000036 & 0.0044639 & 2.000020 & 0.0078308 & 2.000859 & 0.0060766\\
			\hline
			Baseline2 & 1.999971 & 0.0055923 & 2.000246 & 0.0044456 & 1.999739 & 0.0071509 & 1.999992 & 0.0073664\\
			\hline
			Baseline3 & 2.000327 & 0.0054549 & 1.999861 & 0.0048111 & 1.999165 & 0.0078453 & 2.000160 & 0.0063235\\
			\hline
			Baseline4 & 2.000520 & 0.0057238 & 1.999501 & 0.0046540 & 1.999282 & 0.0077363 & 1.999289 & 0.0065290\\
			\hline
			Baseline5 & 1.999919 & 0.0056334 & 2.000053 & 0.0045020 & 2.001443 & 0.0081239 & 2.002397 & 0.0075056\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline1 & 1.999708 & 0.0052113 & 2.000030 & 0.0039910 & 1.999885 & 0.0069901 & 2.001023 & 0.0054369\\
			\hline
			Baseline2 & 1.999734 & 0.0049939 & 2.000314 & 0.0039684 & 2.000046 & 0.0063859 & 1.999684 & 0.0065745\\
			\hline
			Baseline3 & 2.000604 & 0.0048757 & 2.000025 & 0.0042948 & 1.999356 & 0.0069903 & 2.000438 & 0.0056503\\
			\hline
			Baseline4 & 2.000410 & 0.0051162 & 1.999830 & 0.0041504 & 1.999885 & 0.0069068 & 1.999013 & 0.0058261\\
			\hline
			Baseline5 & 1.999830 & 0.0050288 & 1.999830 & 0.0040212 & 2.001592 & 0.0072630 & 2.002090 & 0.0067284\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMA Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline 1 & 1.997416 & 0.3900006 & 2.692152 & 0.5081138 & 35.656866 & 4815.2163162 & 1.667689 & 0.1718689\\
			\hline
			Baseline 2 & 1.686364 & 0.1886132 & 1.679359 & 0.1102722 & 1.474460 & 0.1130752 & 1.534979 & 0.1496990\\
			\hline
			Baseline 3 & 1.572708 & 0.1340966 & 1.743175 & 0.1667602 & 5.345243 & 33.1125597 & 2.313202 & 0.6045124\\
			\hline
			Baseline 4 & 1.640228 & 0.1956211 & 2.046183 & 0.3959176 & 1.327043 & 0.0671306 & -3.008847 & 7.5919771\\
			\hline
			Baseline 5 & 1.590702 & 0.1685914 & 1.842150 & 0.2204031 & 1.488547 & 0.1285866 & 1.868254 & 0.2217948\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRLMAT Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline 1 & 1.398016 & 0.6028170 & 1.421201 & 0.7403641 & 1.049360 & 4.8830065 & 1.401437 & 0.8194417\\
			\hline
			Baseline 2 & 1.406323 & 0.6309170 & 1.442263 & 0.9241642 & 1.369425 & 0.8681724 & 1.449109 & 0.7463529\\
			\hline
			Baseline 3 & 1.408807 & 0.6215050 & 1.358127 & 0.8493445 & 1.375713 & 1.2574743 & 1.436717 & 0.6726785\\
			\hline
			Baseline 4 & 1.387819 & 0.6177830 & 1.390251 & 0.5728970 & 1.596744 & 1.3603545 & 1.325567 & 0.6545989\\
			\hline
			Baseline 5 & 1.387599 & 0.6657501 & 1.386679 & 0.5751290 & 1.359862 & 1.0018556 & 1.538778 & 0.9068372\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline 1 & 2.592974 & 9.040841 & 0.7282379 & 1.286539 & 2.183371 & 7.397726 & 0.7645942 & 1.862201\\
			\hline
			Baseline 2 & 2.210182 & 6.179671 & 0.7485576 & 1.504683 & 1.712691 & 5.071385 & 0.7513498 & 1.560614\\
			\hline
			Baseline 3 & 2.306256 & 6.684895 & 0.7206976 & 1.230857 & 1.718776 & 4.593769 & 0.8374593 & 2.542060\\
			\hline
			Baseline 4 & 2.547665 & 9.843667 & 0.8478792 & 2.387153 & 1.817153 & 5.771054 & 0.8604847 & 2.307269\\
			\hline
			Baseline 5 & 2.224111 & 6.227286 & 0.8080670 & 2.043471 & 1.859343 & 4.996316 & 0.7686838 & 2.308889\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRSIPR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline 1 & 0.9251762 & $<$ 1e-07 & 0.8588266 & $<$ 1e-07 & 0.8386350 & $<$ 1e-07 & 0.7745969 & $<$ 1e-07\\
			\hline
			Baseline 2 & 0.9435412 & $<$ 1e-07 & 0.9081552 & $<$ 1e-07 & 0.8792348 & $<$ 1e-07 & 0.6976176 & $<$ 1e-07\\
			\hline
			Baseline 3 & 0.9694855 & $<$ 1e-07 & 0.9183543 & $<$ 1e-07 & 1.0065755 & $<$ 1e-07 & 0.7129407 & $<$ 1e-07\\
			\hline
			Baseline 4 & 0.9811143 & $<$ 1e-07 & 0.8743186 & $<$ 1e-07 & 0.8665770 & $<$ 1e-07 & 0.9207903 & $<$ 1e-07\\
			\hline
			Baseline 5 & 0.9276948 & $<$ 1e-07 & 0.9107829 & $<$ 1e-07 & 1.1514790 & $<$ 1e-07 & 0.7694052 & $<$ 1e-07\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, LOGLINEAR Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline1 & 1.999840 & 0.0002227 & 1.999761 & 0.0001427 & 1.999820 & 0.0003227 & 2.000443 & 0.0001910\\
			\hline
			Baseline2 & 1.999807 & 0.0002198 & 2.000303 & 0.0001342 & 2.000000 & 0.0002885 & 1.999556 & 0.0002342\\
			\hline
			Baseline3 & 2.000386 & 0.0002191 & 2.000042 & 0.0001598 & 1.998138 & 0.0003215 & 1.999213 & 0.0002316\\
			\hline
			Baseline4 & 2.000039 & 0.0002252 & 1.999710 & 0.0001506 & 1.999217 & 0.0002895 & 1.999525 & 0.0002144\\
			\hline
			Baseline5 & 2.000148 & 0.0002194 & 1.999993 & 0.0001479 & 2.001889 & 0.0003495 & 2.001873 & 0.0002438\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SIRMC Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$,$\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
		\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
			\hline
			Baseline1 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
			\hline
			Baseline2 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
			\hline
			Baseline3 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
			\hline
			Baseline4 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
			\hline
			Baseline5 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07 & 0.9485555 & $<$ 1e-07\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, SEQBAYES Model,
		$X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$, $\beta = 0.06, \gamma = 0.03$}
	\centering
	\begin{footnotesize}
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 1.0000000 & 0.0523424 & 0.8099669 & 0.0471072 & 1.0000000 & 0.0523424 & 1.000000 & 0.0523424\\
		\hline
		Baseline2 & 1.3936378 & 0.0617915 & 1.0000000 & 0.0523424 & 1.0000000 & 0.0523424 & 1.432940 & 0.0626567\\
		\hline
		Baseline3 & 1.4504175 & 0.0630377 & 1.1185890 & 0.0553591 & 1.4331404 & 0.0626611 & 1.000000 & 0.0523424\\
		\hline
		Baseline4 & 1.0000000 & 0.0523424 & 1.3993848 & 0.0619187 & 0.8221337 & 0.0474597 & 1.386606 & 0.0616354\\
		\hline
		Baseline5 & 0.9935711 & 0.0521739 & 1.4389154 & 0.0627872 & 1.3236479 & 0.0602199 & 1.014455 & 0.0527193\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

@@

\begin{table}

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		Baseline1 & 4.421794 & 0.5328029 & 2.280974 & 1.6428406 & 4.442458 & 0.5255489 & 2.609343 & 1.4541148\\
		\hline
		Baseline2 & 1.145541 & 0.0014971 & 1.144060 & 0.0015292 & 1.144700 & 0.0015153 & 1.144115 & 0.0015280\\
		\hline
		Baseline3 & 1.145483 & 0.0014983 & 1.144182 & 0.0015266 & 1.144805 & 0.0015130 & 1.143958 & 0.0015315\\
		\hline
		Baseline4 & 1.145608 & 0.0014956 & 1.144130 & 0.0015277 & 1.144702 & 0.0015152 & 1.143986 & 0.0015309\\
		\hline
		Baseline5 & 1.145550 & 0.0014969 & 1.144154 & 0.0015272 & 1.144711 & 0.0015150 & 1.143915 & 0.0015324\\
		\hline
	\end{tabular}
\end{table}

\begin{table}

>>>>

\message{ !name(draft_v13.tex)}\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
\usepackage{systeme}
%% tables
\usepackage{booktabs}
\usepackage{lscape}
\usepackage[table,xcdraw]{xcolor}
\usepackage{float}

@@

\newcommand{\xxsir}{\ensuremath{11} } % number of SIR methods
\newcommand{\wxxsir}{11 }
\newcommand{\Wxxsir}{11 }

@@

\message{ !name(draft_v13.tex) !offset(-3) }

@@

In terms of data from the SIR model, incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.  This method assumes that we have some prior knowledge about $\gamma$, the recovery rate.  Thus we use as our estimate,
\begin{align*}
\hat{\rr} &= \textnormal{IPR}(t) \cdot \frac{1}{\gamma}
\end{align*}

@@

\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection. Then, we have
\begin{align}\label{eq:lotka}
\rr = \exp{(r \omega)}.
\end{align}
Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.

\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
\begin{align*}
\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
\end{align*}
\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.  They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.  Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
\begin{align*}
  M(-r) = \frac{\gamma}{\gamma + r},
\end{align*}
and the resulting estimate of $\rr$ is
\begin{align}\label{eq:r0-eg}
  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
\end{align}
where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.  The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.  We may estimate $\hat{r}$ using ordinary linear regression, namely,
\begin{align*}
  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
  \epsilon_t &\sim N(0, \sigma^2),
\end{align*}
which results in
\begin{align}\label{eq:exp-growth}
  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
\end{align}

For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.  Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.  We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used. The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.  This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.

An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.  Another advantage is its flexibility.  Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).  However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.  Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.

\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.  They model the incidence counts during the $t$th interval, $J(t),$ as
\begin{align*}
  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
\end{align*}
where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
\begin{align*}
  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
\end{align*}
The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.  Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.

Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.  Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.  A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.  Also, in this method we must select the interval size of the incidence counts.

@@

We compare the performance of the SIR models on simulated data from both the SIR as well as other models. In all of these simulations, we first generate data from the models under known conditions, then we add on errors drawn from known distributions variances to each time point,
\begin{align}\label{eq:sim-models}
  X_{model}(t) &= f(t) + \epsilon_{X,t} \\
  Y_{model}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
  Z_{model}(t) &= N - X(t) - Y(t)\nonumber 
\end{align}
We adjust the recovered compartment $Z$ so the total population is constant. 
We simulate with two different types of error distributions, Gaussian and autoregressive,

@@

\noindent \textbf{Gaussian-PAVA (Norm-M) ($X_{GP}$, $Y_{GP}$, $Z_{GP}$)}
\begin{align*}
 X_{GP}(t) &= \textnormal{PAVA}(X_G(t)) \\
  Y_{GP}(t) &= N - X_{GP}(t) - Z_{GP}(t) \\
  Z_{GP}(t) &= \textnormal{PAVA}(Z_G(t)) \\
\end{align*}
\textbf{Autoregressive-PAVA (AR-M) ($X_{ARP}$, $Y_{ARP}$, $Z_{ARP}$)}
\begin{align*}
  X_{ARP}(t) &= \textnormal{PAVA}(X_{AR}(t)) \\
  Y_{ARP}(t) &= N - X_{ARP}(t) - Z_{ARP}(t) \\
  Z_{ARP}(t) &= \textnormal{PAVA}(Z_{AR}(t)) 
\end{align*}
This gives us a total of four error structures for each set of conditions: Gaussian (Norm), Gaussian monotone (Norm-M), autoregressive (AR), and autoregressive monotone (AR-M) for a given set of parameters $(\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N)$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals, the standard deviation for both the susceptible and infectious compartments, and the total population, respectively.

@@

We first simulate a ``baseline'' data set from the SIR model under the following conditions:

\textbf{Baseline Dataset}
\begin{center}

@@

	Total population $N$: 100000

        Error: Gaussian.

@@

This baseline data set

% We then generate different SIR data sets, changing the one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison. The different data sets generated are shown in the following table:

% \begin{table}[h]
% 	\centering
% 	\resizebox{\textwidth}{!}{
% 		\begin{tabular}{@{}lll@{}}
% 			\toprule
%                   \textbf{Parameter Values ($\beta, \gamma$)}        & \textbf{Starting Compartment Sizes ($X_0, Y_0$)}            & \textbf{Variance of Added Errors ($\sigma_X, \sigma_Y$ )}       \\ \midrule
%                   (0.06, 0.001)            & (99000, 1000)                           & (50, 2)                     \\
%                   (0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
%                   (0.06,.03) & (99950, 50) & (100, 5)\\
%                   (0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
%                   (0.06, 0.24) & (99999, 1) & (10000, 500)                                      \\ \bottomrule
% 		\end{tabular}
% 	}
% 	\caption{Table of data sets.}
% 	\label{tab:datasets}
% \end{table}

@@

\section{Results}\label{sec:results}
There are numerous aspects to consider when using a method to estimate $\rr$.  Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available (time $T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, and  the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X, \sigma_Y$).  We first examine our baseline data, and then we examine each of these aspects individually.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{images/BaseBase.jpeg}
  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The red line is the true $\rr$ value.}
  \end{figure}

\begin{table}[H]	
	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		SIRLS & 1.9999711 & 0.0055923\\
		\hline
		SIRREPAR & 1.9997340 & 0.0049939\\
		\hline
		SIRLMA & 1.6863640 & 0.1886132\\
		\hline
		SIRLMAT & 1.4063227 & 0.6309170\\
		\hline
		SIRIPR & 4.4203632 & 12.3593415\\
		\hline
		SIRSIPR & 1.8870824 & $<$ 1e-07 \\
		\hline
		LOGLINEAR & 1.9998072 & 0.0002198\\
		\hline
		SIRMC & 0.9485555 &  $<$ 1e-07 \\
		\hline
		SEQBAYES & 1.6495023 & 0.0672249\\
		\hline
	\end{tabular}
        \caption{Baseline Data $(\beta=.06, \gamma=.03, O=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline}
\end{table}

@@

\includegraphics[scale=0.5]{images/AR.jpeg}
\begin{table}[H]

@@

\includegraphics[scale=0.5]{images/parchange.jpeg}

@@

\includegraphics[scale=0.5]{images/time.jpeg}

@@

\includegraphics[scale=0.5]{images/start.jpeg}

@@

\includegraphics[scale=0.5]{images/var.jpeg}

@@

  \begin{center}
\includegraphics[scale=0.5]{images/other.jpeg}

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

	\caption{\label{tab:}$\rr$ Estimates and Std. Errs, EXPGROWTH Model,
		$\beta = 0.06, \gamma = 0.03$, $X_0 = 99950, Y_0 = 50$, $\sigma_X = 100, \sigma_Y = 5$}
	\centering
	\begin{tabular}[t]{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Norm Est & Norm SE & Norm-M Est & Norm-M SE & AR Est & AR SE & AR-M Est & AR-M SE\\
		\hline
		$\gamma$ = 0.001 & 1.141015 & 0.0001503 & 1.1349235 & 0.0001645 & 1.139834 & 0.0001529 & 1.1348807 & 0.0001646\\
		\hline
		$\gamma$ = 0.04 & 1.133308 & 0.0001686 & 1.1310087 & 0.0001744 & 1.132539 & 0.0001705 & 1.1313483 & 0.0001735\\
		\hline
		$\gamma$ = 0.06 & 1.093381 & 0.0003053 & 0.9905347 & 0.0013816 & 1.085491 & 0.0003433 & 0.9899559 & 0.0013931\\
		\hline
		$\gamma$ = 0.24 & 1.092493 & 0.0003093 & 0.9206087 & 0.0036121 & 1.085583 & 0.0003428 & 0.8860333 & 0.0055673\\
		\hline
	\end{tabular}
\end{table}
